# PA1_template

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading the dataset

This is a report of analysis of the "activity.csv" dataset.

```{r read}
data <- read.csv("activity.csv")
head(data)
```

## Processing the dataset

This is to change the format of the date from character format to Date format.

```{r date}
data$date <- as.Date(data$date,format="%Y-%m-%d")
head(data)
```

## Plotting histogram of total number of steps per day

aggregate function is used to sun the number of steps in each given day.

```{r hist}
steps_in_a_day <- aggregate(steps~date,data,sum)
with(steps_in_a_day,hist(steps,breaks = 10))
```

## Mean and Median of steps in a day

```{r mean_median}
aggregate(steps~date,data,mean)
aggregate(steps~date,data,median)
```

Median of the steps on all days is 0 because for most of the 5 minutes interval, the person is at rest, i.e number of steps is 0.

## Mean by interval in the day

aggregate function is used again to calculate average number of steps in a given time interval. The interval is given the form of "2150" as in 21 hours and 50 minutes. Hence, I used integer operations to convert this into minutes as 21*60+50 = 1310 minutes. Total number of minutes in a day is 1440.

```{r time_series}
steps_average_by_interval <- aggregate(steps~interval,data,mean)
steps_average_by_interval$minutes <- (steps_average_by_interval$interval%/%100)*60 +steps_average_by_interval$interval%%100
with(steps_average_by_interval,plot(minutes,steps,type="l"))
```

## Time interval with maximum average number of steps

```{r max_interval}
max_interval <- which(steps_average_by_interval$steps == max(steps_average_by_interval$steps))
steps_average_by_interval$interval[max_interval]
```

## Total missing values in the dataset

```{r total_na}
sum(is.na(data))
```

## Imputing values to create a new dataset

ifelse() function has been used. If the value is missig, the steps has been set as average number of steps in that particular time interval.

```{r new_dataset}
test <- data
test$steps <- ifelse(is.na(test$steps),steps_average_by_interval$steps[steps_average_by_interval$interval==test$interval] , test$steps)
```

## Plots and analysis on the new dataset

```{r new_dataset_analysis}
steps_in_a_day_imputed <- aggregate(steps~date,test,sum)
with(steps_in_a_day_imputed,hist(steps,breaks = 10))

aggregate(steps~date,test,mean)
aggregate(steps~date,test,median)
```

Mean of steps with days having NA values increased. Median of steps for a day remained at 0 except for 1st days which shows a non-zero value because most of the data on that day is imputed based on average.

## Creating a new column of factor varible weekday and weekend

```{r weekday}
test$weekday <- weekdays(test$date)
test$weekday <- test$weekday == "Saturday" | test$weekday == "Sunday"
test$weekday <- factor(test$weekday,labels = c("weekday","weekend"))
```

## Plotting two time series plots of weekdays and weekend

```{r plot_time_series_by_weekday}
steps_average_by_interval_weekdays <- aggregate(steps~interval,subset(test,weekday=="weekday"),mean)
steps_average_by_interval_weekend <- aggregate(steps~interval,subset(test,weekday=="weekend"),mean)

par(mfrow=c(1,2))

plot(steps_average_by_interval$minutes,steps_average_by_interval_weekdays$steps,type="l",ylim = c(0,250))
plot(steps_average_by_interval$minutes,steps_average_by_interval_weekend$steps,type="l",ylim = c(0,250))
```
